---
title: "Overmodification Experiment 2: Processing Script (Version 2.0)"
format: 
  html:
    toc: true
    toc-location: right
    toc-depth: 2 
embed-resources: true
editor: visual
execute: 
  warning: false
---

A processing script for experiment two of the overmodification project

*Set up computing environment*

```{r setup}
#| echo: true 
#| output: false

library(tidyverse)
library(jsonlite)
library(brms)
library(tidybayes)
library(BayesFactor)
library(ggthemr)
ggthemr("pale")


sand = c("#e0cda9")
dust = c("#9E8472")
forest = c("#a0ae9b")
shallowWaters = c("#a4b8c1")

darkSand = c("#5C4033")
darkDust = c("#1b0000")
darkForest = c("#06402B")
darkShallowWaters = c("#2f4858")

earthShades = c("#b2988e", "#b37a5f", "#d4b88a", "#a0ae9b", "#8d9da8")

earthShadesDark = c("#554740", "#553929", "#68553d", "#06402B", "#2f4858")

fourPalette = c("#7a6b6e", "#8d9da8", "#e0cda9", "#a8a89e")

fourPaletteDark = c("#1b0000", "#2f4858", "#5C4033", "#06402B")

threePalette = c("#8d9da8", "#a8a89e", "#e0cda9")

threePaletteDark = c("#2f4858", "#06402B", "#5C4033")

data <- read_csv("data/rawData.csv")

finishStatusOrient <- read_csv("data/orientationFinishStatuses.csv")
finishStatusColorLoFreq <- read_csv("data/colorLoFreqFinishStatuses.csv")
finishStatusColorHiFreq <- read_csv("data/colorHiFreqFinishStatuses.csv")


#for ease of col name use
colnames(data)<- c("serverCode", "column1", "column2", "string")
```

###### 

*Removing participants who did not complete the experiment (in-line with ethics approval), and experimenters debug link*

```{r}
# experimenters debugging
data <- data %>%   
  select(-column1) %>%   
  filter(!grepl('debug', serverCode)) %>%    
  separate(
    serverCode, 
    into = c("Participant id", NA), 
    sep = ":", 
    remove=FALSE
    )

# drop outs
finishStatus <- bind_rows(finishStatusOrient, finishStatusColorLoFreq, finishStatusColorHiFreq) 

data <- finishStatus %>%
  select("Participant id", "Status") %>% 
  left_join(data) %>% 
  filter(Status == "APPROVED") %>% 
  select(-`Participant id`, -Status) 
```

## Transforming JSON strings to nice tibbles

*Turn server codes into participants with id's, and filter out unnecessary columns.*

```{r}
participantData <- data %>%   
  separate(serverCode, 
           into = c("participant", NA), 
           sep = ":", 
           remove=FALSE) %>% 
  group_by(participant) %>% 
  mutate(participant = if(n() > 1) { 
    sprintf("participant%03d",                   
            cur_group_id())     
    }) %>% 
  ungroup() %>% 
  select(participant, 
         everything(), 
         -serverCode) %>% 
  arrange(participant) %>% 
  select(-column2)
```

*Check our total N*

```{r}
usablePreEverything <- participantData$participant %>% 
  unique()

message('Everything including rejections and returns: ', length(usablePreEverything), ' people.')
```

As expected, 60 participants.

###### 

*Filter in relevant data*

```{r}
data <- participantData %>% 
  filter(str_detect(tolower(string), "setsize") | 
         str_detect(tolower(string), "condition"))
```

*Parse, and also removing full trial here, to take the data we need.*

```{r}
parseddata <- data %>% 
  mutate(inx = 1:n()) %>% 
  rowwise() %>% 
  mutate(trialdata = lapply(string, fromJSON)) %>%
  unnest(trialdata) %>% 
  mutate(label = names(trialdata)) %>% 
  spread(label, trialdata) %>% 
  # remove full trial
  select(-fullTrial) %>% 
  unnest(where(is.list)) %>% 
  select(-string, -inx) 
```

*Demarcate participants in low freq condition*

```{r}
# Assuming you have a participant identifier column like "participantID"
# First, identify which participants have the "nonFreqColorTerms" condition
nonFreqParticipants <- parseddata %>%
  filter(condition == "nonFreqColorTerms") %>%
  distinct(participant) %>%
  pull(participant)

# Now update targetColor for all rows belonging to those participants
parseddata <- parseddata %>%
  mutate(
    targetColor = case_when(
      participant %in% nonFreqParticipants & targetColor == "blue" ~ "teal",
      participant %in% nonFreqParticipants & targetColor == "green" ~ "jade",
      TRUE ~ targetColor
    )
  )
```

###### 

*Make low freq redundant values change*

```{r}
# First, identify participants who have teal or jade target colors
participants_with_low_freq_colors <- parseddata %>%
  filter(targetColor %in% c("teal", "jade")) %>%
  distinct(participant) %>%
  pull(participant)

# Update redundant column for all rows belonging to those participants
parseddata <- parseddata %>%
  mutate(
    redundant = case_when(
      participant %in% participants_with_low_freq_colors ~ "colorredundantlowfreq",
      TRUE ~ redundant
    )
  )

# remove condition column 
parseddata <- parseddata %>% 
  select(-condition) 
```

*Save the dataframe*

```{r}
write_csv(parseddata, "data/parsedDataFrame.csv")
```

*Check our total N recruited after filtering in relevant data*

```{r}
usableAfterFirstFilter <- parseddata$participant %>% 
  unique()

message('Total recruited is: ', length(usableAfterFirstFilter), ' people.')
```

N still at 60, as it should be.

###### 

###### 

## Participant Removal

There is just one applicable criteria here: accuracy. Whether the noun used matches the actual shape.

*Filter in crit data*

```{r}
parsedCrit <- parseddata %>% 
  filter(
    trialType == "crit"
  )
```

*Mutate accuracy*

```{r}
accuracyParsedCrit <- parsedCrit %>%
  mutate(accurate = str_detect(response, fixed(targetShape)))
```

*Check for unacceptable participants (\>50% as pre-registered)*

```{r}
participantAccuracy <- accuracyParsedCrit %>%
  group_by(participant) %>%
  summarise(
    total_trials = n(),
    false_trials = sum(accurate == FALSE, na.rm = TRUE),
    proportion_false = false_trials / total_trials
  ) %>%
  filter(proportion_false > 0.5) %>% view()

```

No removals! Woohoo!

*Compute n-trials with inaccuracies*

```{r}
inaccurateTrials <- accuracyParsedCrit %>% 
  filter(accurate == FALSE)

# n trials removed
nrow(inaccurateTrials)

# % trials removed 
nrow(inaccurateTrials) / nrow(accuracyParsedCrit)

```

44 trials removed for inaccuracy. 0.2% of trials.

*Filter out the rest of inaccuracies via individual trials, and remove accurate because we don't need it.*

```{r}
parsedcrit <- accuracyParsedCrit %>% 
  filter(accurate == TRUE) %>% 
  select(-accurate) 
```

###### 

#### Taking stock

```{r}
usableAfterExclusions <- parsedcrit$participant %>% 
  unique()

usableTrials <- nrow(parsedcrit)

message('Total participants analysed: ', length(usableAfterFirstFilter), ' people.')

message('Total number of critical trials: ', usableTrials)
```

And that's it! Hard to believe after the hell of the last experiment.

###### 

###### 

## Building the Final Tibble

*Create overinformativeness*

```{r}
nouns <- c("circle", "square", "hexagon", "pentagon")
```

*Create overinformativeness column as any response with more than the nouns*

```{r}
process_response <- function(response) {
  words <- str_split(tolower(response), "\\s+")[[1]]
  nonNounCount <- sum(!words %in% nouns & words != "")
  return(nonNounCount)
}

tibble <- parsedcrit %>%
  mutate(across(where(is.character), tolower)) %>%
  rowwise() %>%
  mutate(
    nonNounCount = process_response(response),
    overinformative = nonNounCount >= 1
  ) %>%
  select(-nonNounCount) %>%
  ungroup()
```

*Manual Check*

```{r}
tibble %>% 
  select(overinformative, response) 
```

*Rename for plotting and working with df*

```{r}
tibble <- tibble %>%
    mutate(setSize = recode(setSize, setsize3 = 'setsize03')) %>% 
  mutate(setSize = recode(setSize, setsize6 = 'setsize06')) %>% 
  mutate(setSize = recode(setSize, setsize9 = 'setsize09'))
```

###### 

###### 

## Whaddowegot?

*Colour vs. Orientation Redundant*

```{r}
tibble %>% 
  filter(trialType == "crit") %>%
  mutate(od = ifelse(overinformative, 1, 0)) %>% 
  ggplot(
    aes(x = redundant, 
        y = od, 
        fill = redundant,
        color = redundant)
    ) + 
  stat_summary(fun.y = mean, geom="bar", 
               alpha = 0.9, show.legend=FALSE, 
               color = NA) + 
  geom_jitter(
      stat = "summary", aes(group = participant),
      fun = mean, alpha = 0.25, size = 2.8, 
      show.legend = FALSE, width = 0.2
      ) +
  stat_summary(fun.data = mean_cl_boot, 
               geom="linerange", size = 1, 
               color = "grey25") +
  coord_cartesian(ylim = c(0, 1)) +
  ylab("Proportion Overinformative") + 
  xlab("Redundant Attribute") +
  theme(
    text = element_text(size = 18, 
                        family = "serif",
                        color = "black"), 
    panel.border = element_rect(color = "gray80", 
                                fill = NA, 
                                size = 1),
    panel.grid = element_blank()
    ) +
 scale_x_discrete(
    limits = c("orientredundant", "colorredundantlowfreq", "colorredundant"),
  
    labels = c("Orientation",  "Colour \n LF", "Colour \n HF")
 ) + 
  scale_fill_manual(values = threePalette) +
  scale_color_manual(values = threePaletteDark)
```

###### 

*Comparison of set sizes, faceted by redundant attribute*

```{r}
labels <- c(
  "colorredundant" = "HF Colour Redundant",
  "colorredundantlowfreq" = "LF Colour Redundant",
  "orientredundant" = "Orientation Redundant"
)

tibble %>% 
  mutate(od = ifelse(overinformative, 1, 0)) %>% 
  filter(trialType == "crit") %>% 
  ggplot(
    aes(x = setSize, 
        y = od, 
        fill = setSize,
        color = setSize)
    ) + 
  facet_wrap(~redundant,
             labeller = as_labeller(labels)
               ) +
  stat_summary(fun.y = mean, geom="bar", 
               alpha = 0.9, show.legend=FALSE, 
               color = NA) + 
  geom_jitter(
      stat = "summary", aes(group = participant),
      fun = mean, alpha = 0.4, size = 2.8, 
      show.legend = FALSE, width = 0.1
      ) +
  stat_summary(fun.data = mean_cl_boot, 
               geom="linerange", size = 1, 
               color = "grey25") +
  coord_cartesian(ylim = c(0, 1)) +
  ylab("Proportion Overinformative") + 
  xlab("Display Set Size") +
  theme(
    text = element_text(size = 18, 
                        family = "serif",
                        color = "black"), 
    panel.border = element_rect(color = "gray80", 
                                fill = NA, 
                                size = 0.7),
    panel.grid = element_blank()
    ) +
  scale_fill_manual(values = fourPalette
                    ) + 
  scale_color_manual(values = fourPaletteDark
                     ) +
  scale_x_discrete(
    labels = c("Three", "Six", "Nine", "Sixteen")
    ) 
```

###### 

*Comparison of n-distractors, faceted by redundant attribute*

```{r}
labels <- c(
  "colorredundant" = "HF Colour Redundant",
  "colorredundantlowfreq" = "LF Colour Redundant",
  "orientredundant" = "Orientation Redundant"
)

tibble %>% 
  mutate(od = ifelse(overinformative, 1, 0)) %>% 
  filter(trialType == "crit") %>% 
  ggplot(
    aes(x = nDistract, 
        y = od, 
        fill = nDistract,
        color = nDistract)
    ) + 
  facet_wrap(~redundant,
             labeller = as_labeller(labels)
               ) +
  stat_summary(fun.y = mean, geom="bar", 
               alpha = 0.9, show.legend=FALSE, 
               color = NA) + 
  geom_jitter(
      stat = "summary", aes(group = participant),
      fun = mean, alpha = 0.4, size = 2.8, 
      show.legend = FALSE, width = 0.1
      ) +
  stat_summary(fun.data = mean_cl_boot, 
               geom="linerange", size = 1, 
               color = "grey25") +
  coord_cartesian(ylim = c(0, 1)) +
  ylab("Proportion Overinformative") + 
  xlab("Number of Distractors") +
  theme(
    text = element_text(size = 18, 
                        family = "serif",
                        color = "black"), 
    panel.border = element_rect(color = "gray80", 
                                fill = NA, 
                                size = 0.7),
    panel.grid = element_blank()
    ) +
  scale_fill_manual(values = earthShades) +
  scale_color_manual(values = earthShadesDark) + 
  scale_x_discrete(
    labels = c("None", "One", "Two", "Three", "Four")
    ) 
```

###### 

###### 

## Inferential Statistics

### Null

*Set up*

```{r}
#| echo: true 
#| output: false

null <- brm(
  formula = overinformative ~ 1,
  data = tibble,
  family = bernoulli(link = "logit"), #binary logistic regression
  warmup = 5000,  #burn-in   
  iter = 10000,  #samples    
  chains = 4,  #n chains    
  cores=4, # cpu cores running in parallel  
  #prior = prior("normal(0, 2)", class = "b"),
  seed = 3000, #seed so we can return
  save_pars = save_pars(all = TRUE)
  )
```

*Summary*

```{r}
summary(null)

# chain convergence
mcmc_plot(null,
         type = "trace",
         prob = 0.95)

# point estimate visualisation
mcmc_plot(null,
         type = "areas",
         prob = 0.95)
```

 

### Full Model

*Set-up*

```{r}
#| echo: true 
#| output: false

full <- brm(
  formula = overinformative ~ 1 + redundant + nDistract + setSize,
  data = tibble,
  family = bernoulli(link = "logit"), #binary logistic regression
  warmup = 5000,  #burn-in   
  iter = 10000,  #samples    
  chains = 4,  #n chains    
  cores=4, # cpu cores running in parallel  
  prior = prior("normal(0, 2)", class = "b"),
  seed = 3000, #seed so we can return
  save_pars = save_pars(all = TRUE)
  )
```

*Summary*

```{r}
summary(full)

# chain convergence
mcmc_plot(full,
         type = "trace",
         prob = 0.95)

# point estimate visualisation
mcmc_plot(full,
         type = "areas",
         prob = 0.95)
```

###### 

### Setsize + Attribute-redundant

*Set up*

```{r}
#| echo: true 
#| output: false

SetAtt  <- brm(
  formula = overinformative ~ 1 + redundant + setSize,
  data = tibble,
  family = bernoulli(link = "logit"), #binary logistic regression
  warmup = 5000,  #burn-in   
  iter = 10000,  #samples    
  chains = 4,  #n chains    
  cores=4, # cpu cores running in parallel  
  prior = prior("normal(0, 2)", class = "b"),
  seed = 3000, #seed so we can return
  save_pars = save_pars(all = TRUE)
  )
```

*Summary*

```{r}
summary(SetAtt)

# chain convergence
mcmc_plot(SetAtt,
         type = "trace",
         prob = 0.95)

# point estimate visualisation
mcmc_plot(SetAtt,
         type = "areas",
         prob = 0.95)
```

###### 

### N-distractors + Attribute-redundant

*Set up*

```{r}
#| echo: true 
#| output: false

DistAtt  <- brm(
  formula = overinformative ~ 1 + redundant + nDistract,
  data = tibble,
  family = bernoulli(link = "logit"), #binary logistic regression
  warmup = 5000,  #burn-in   
  iter = 10000,  #samples    
  chains = 4,  #n chains    
  cores=4, # cpu cores running in parallel  
  prior = prior("normal(0, 2)", class = "b"),
  seed = 3000, #seed so we can return
  save_pars = save_pars(all = TRUE)
  )
```

*Summary*

```{r}
summary(DistAtt)

# chain convergence
mcmc_plot(DistAtt,
         type = "trace",
         prob = 0.95)

# point estimate visualisation
mcmc_plot(DistAtt,
         type = "areas",
         prob = 0.95)
```

###### 

### Set Size

*Set up*

```{r}
#| echo: true 
#| output: false

setSize  <- brm(
  formula = overinformative ~ 1 + setSize,
  data = tibble,
  family = bernoulli(link = "logit"), #binary logistic regression
  warmup = 5000,  #burn-in   
  iter = 10000,  #samples    
  chains = 4,  #n chains    
  cores=4, # cpu cores running in parallel  
  prior = prior("normal(0, 2)", class = "b"),
  seed = 3000, #seed so we can return
  save_pars = save_pars(all = TRUE)
  )
```

*Summary*

```{r}
summary(setSize)

# chain convergence
mcmc_plot(setSize,
         type = "trace",
         prob = 0.95)

# point estimate visualisation
mcmc_plot(setSize,
         type = "areas",
         prob = 0.95)
```

###### 

### N distractors

*Set up*

```{r}
#| echo: true 
#| output: false

nDistractor  <- brm(
  formula = overinformative ~ 1 + nDistract,
  data = tibble,
  family = bernoulli(link = "logit"), #binary logistic regression
  warmup = 5000,  #burn-in   
  iter = 10000,  #samples    
  chains = 4,  #n chains    
  cores=4, # cpu cores running in parallel  
  prior = prior("normal(0, 2)", class = "b"),
  seed = 3000, #seed so we can return
  save_pars = save_pars(all = TRUE)
  )
```

*Summary*

```{r}
summary(nDistractor)

# chain convergence
mcmc_plot(nDistractor,
         type = "trace",
         prob = 0.95)

# point estimate visualisation
mcmc_plot(nDistractor,
         type = "areas",
         prob = 0.95)
```

### Attribute Redundant

*Set up*

```{r}
#| echo: true 
#| output: false

attribute  <- brm(
  formula = overinformative ~ 1 + redundant, 
  data = tibble, 
  family = bernoulli(link = "logit"), #binary logistic  regression 
  warmup = 5000,  #burn-in   
  iter = 10000,  #samples    
  chains = 4,  #n chains    
  cores=4, # cpu cores running in parallel  
  prior = prior("normal(0, 2)", class = "b"), 
  seed = 3000, #seed for determinism 
  save_pars = save_pars(all = TRUE) 
  )
```

*Summary*

```{r}
summary(attribute)

# chain convergence
mcmc_plot(attribute,
         type = "trace",
         prob = 0.95)

# point estimate visualisation
mcmc_plot(attribute,
         type = "areas",
         prob = 0.95)
```

###### 

###### 

### Model Comparison

```{r}
bayes_factor(full, null, log = TRUE)
bayes_factor(DistAtt, null, log = TRUE)
bayes_factor(SetAtt, null, log = TRUE)
bayes_factor(setSize, null, log = TRUE)
bayes_factor(nDistractor, null, log = TRUE)
bayes_factor(attribute, null, log = TRUE)
```