---
title: "Staircase Logit Plots"
format: 
  html:
    toc: true
    toc-location: right
    toc-depth: 2 
embed-resources: true
editor: visual
execute: 
  warning: false
---

```{r}
#| echo: true 
#| output: false

library(tidyverse)
library(here)
library(ggthemr)
ggthemr("pale")

light <- c("#DCBCBC")
mid <- c("#B97C7C")
dark <- c("#8F2727")
darker <- c("#7C0000")

lightblue <- c("#BCC8DC")    
midblue <- c("#7C97B9")      
darkblue <- c("#27548F")     
darkerblue <- c("#00247C")

# for extracting border colour of ggthemr 
current_theme <- theme_get()
axis_color <- current_theme$axis.line$colour

# my custom theme
my_theme <- function() {
  theme(
    panel.grid.major = element_line(
      size = 0.2,        
      linetype = "solid"),
    text = element_text(family = "serif", size = 18),
    panel.border = element_rect(color = axis_color, 
                                fill = NA, 
                                linewidth = 1.1),
    axis.line = element_blank()
  )
}

setwd(here("experimentOne", "data", "staircasePlots"))

colour <- read_csv(file = "colorStaircases.csv")
material <- read_csv(file = "materialStaircases.csv")
```

 

## Plotting

 

### Yellow to Orange Staircase Plot

```{r}
steps_metal <- material |>
  mutate(stim = ifelse(dimension == "wood", 10 - step, step),
         correct = ifelse(response == "wood", 1, 0)) |>
  group_by(stim) |> 
   summarise(
    n = n(),
    sum_correct = sum(correct),
    prop_correct = sum_correct / n
  )

steps_yo_binned <- colour |>
  filter(dimension %in% c("yellow", "orange")) |> 
  mutate(stim = ifelse(dimension == "yellow", 50 - step, step),
         correct = ifelse(response == "yellow", 1, 0)) |> 
  mutate(stim = 5 * .bincode(stim, c(1:10) * 5))

steps_yo <- colour |>
  filter(dimension %in% c("yellow", "orange")) |> 
  mutate(stim = ifelse(dimension == "yellow", 50 - step, step),
         correct = ifelse(response == "yellow", 1, 0))
```

```{r}
# Get unique participants
participants <- unique(steps_yo_binned$participant)

# Loop through each participant and create a plot
for (p in participants) {
  # Filter data for current participant
  participant_data <- steps_yo_binned[steps_yo_binned$participant == p, ]
  
  # Create plot
  plot <- ggplot(
    data = participant_data,
    aes(x = stim, y = correct)
  ) +
    geom_point(
      col = "gray30",
      size = 2,
      alpha = 0.5
    ) +
    stat_smooth(
      method = "glm",
      se = FALSE,
      method.args = list(family = binomial),
      col = "lightblue",
      linewidth = 1.5
    ) +
    theme_classic() +
    labs(
      x = "Orange-to-Yellow Staircase Step Number",
      y = "Probability of Label \"yeOlow\"",
      title = paste("Participant", p)
    ) +
    scale_x_continuous(breaks = 0:50) +
    scale_y_continuous(
      "Probability of label \"yellow\"",
      sec.axis = sec_axis(
        ~.,
        name = "Probability of label \"orange\""
      )
    )
  
  # Print the plot
  print(plot)
}
```

```{r}
yo + scale_y_continuous(
  "Probability of label \'yellow\'", 
  sec.axis=sec_axis(
    ~ 1 - ., 
    name="Probability of label \'orange\'"
    )
  )
```

&nsbp;

### Green to Blue Staircase Plot

```{r}
steps_gb_binned <- colour |>
  filter(dimension %in% c("green", "blue")) |> 
  mutate(stim = ifelse(dimension == "green", 50 - step, step),
         correct = ifelse(response == "green", 1, 0)) |> 
  mutate(stim = 5 * .bincode(stim, c(1:10) * 5))

steps_gb <- colour |>
  filter(dimension %in% c("green", "blue")) |> 
  mutate(stim = ifelse(dimension == "green", 50 - step, step),
         correct = ifelse(response == "green", 1, 0))
```

```{r}
gb <- ggplot(
  data = steps_gb_binned, 
  aes(x = stim, y = correct)
) +
# individual participant curves
  stat_smooth(
    aes(group = participant),  
    method = "glm", 
    se = FALSE, 
    method.args = list(family = binomial), 
    col = lightblue,  
    alpha = 0.001,
    size = 0.5
  ) +
  stat_smooth(
    data = steps_gb, 
    aes(x=stim, y = correct),
    method="glm", 
    se=FALSE, 
    method.args = list(family=binomial), 
    col=midblue,
    size = 1.7
    ) +
    stat_summary(fun.data = mean_cl_boot, geom="linerange", color = midblue, alpha = 1, size = 1.8) +
  stat_summary(fun = mean, geom="point", color = darkblue, size = 2.5) +
  my_theme() + 
  labs(
    x = "Blue-to-Green Staircase Step Number",
    y = "Probability of Label \"green\"", 
  )
```

```{r}
gb + scale_y_continuous(
  "Probability of label \'green\'", 
  sec.axis=sec_axis(
    ~ 1 - ., 
    name="Probability of label \'blue\'"
    )
  )
```

 

### Extracting the separable material categories from the list information.

I had to get the list information by going to the original experiment and checking the lists. List 1 is orange-yellow/wood-glass, 2 is orange-yellow/cardboard-metal, 3 is blue-green/wood-glass, and 4 is blue-green/cardboard-metal.

 

### Wood to Glass Staircase Plot

```{r}
steps_wood_glass <- material |> 
  filter(list %in% c("list1", "list3")) |> 
  mutate(stim = ifelse(dimension == "metal", 10 - step, step),
         correct = ifelse(response == "metal", 1, 0))
```

```{r}
wg <- ggplot(
  data = steps_wood_glass, 
  aes(x = stim, y = correct)
) +
    # individual participant curves
  stat_smooth(
    aes(group = participant),  
    method = "glm", 
    se = FALSE, 
    method.args = list(family = binomial), 
    col = light,  
    alpha = 0.01,
    size = 0.5
  ) +
  stat_smooth(
    data = steps_wood_glass, 
    aes(x=stim, y = correct),
    method="glm", 
    se=FALSE, 
    method.args = list(family=binomial), 
    col=mid,
    size = 1.7
    ) +
    stat_summary(fun.data = mean_cl_boot, geom="linerange", color = mid, alpha = 1, size = 1.8) +
  stat_summary(fun = mean, geom="point", color = dark, size = 2.5) +
  my_theme() + 
  labs(
    x = "Wood-to-Glass Staircase Step Number",
    y = "Probability of label \"metal\"", 
  )
```

```{r}
wg + scale_y_continuous(
  "Probability of label \'metal\'", 
  sec.axis=sec_axis(
    ~ 1 - ., 
    name="Probability of label \'wood\'"
    )
  )
```

  

```{r}
steps_cardboard_metal <- material |> 
  filter(list %in% c("list2", "list4")) |> 
  mutate(stim = ifelse(dimension == "metal", 10 - step, step),
         correct = ifelse(response == "metal", 1, 0))
```

```{r}
cm <- ggplot(
  data = steps_cardboard_metal, 
  aes(x = stim, y = correct)
) +
  # individual participant curves
  stat_smooth(
    aes(group = participant),  
    method = "glm", 
    se = FALSE, 
    method.args = list(family = binomial), 
    col = light,  
    alpha = 0.01,
    size = 0.5
  ) +
  # Overall logit curve
  stat_smooth(
    method = "glm", 
    se = FALSE, 
    method.args = list(family = binomial), 
    col = mid,
    size = 1.7
  ) +
    stat_summary(fun.data = mean_cl_boot, geom="linerange", color = mid, alpha = 1, size = 1.8) +
  stat_summary(fun = mean, geom="point", color = dark, size = 2.5) +
  my_theme() + 
  labs(
    x = "Cardboard-to-Metal Staircase Step Number",
    y = "Probability of Label \"Metal\"", 
  )
```

```{r}
cm + scale_y_continuous(
  "Probability of label \'metal\'", 
  sec.axis=sec_axis(
    ~ 1 - ., 
    name="Probability of label \'wood\'"
    )
  )
```
