---
title: "Experiment One: Analysis"
format: 
  html:
    toc: true
    toc-location: right
    toc-depth: 2 
embed-resources: true
editor: visual
execute: 
  warning: false
---

Analysis script for Experiment One. I'll first conduct main analysis, then move on to the supplementary analyses outlined in the pre-registration.

###### 

## Set up Computing Environment

```{r}
#| echo: true 
#| output: false

library(tidyverse)
library(jsonlite)
library(brms)
library(tidybayes)
library(BayesFactor)
library(ggthemr)
ggthemr("pale")


sand = c("#e0cda9")
dust = c("#9E8472")
forest = c("#a0ae9b")

darkSand = c("#5C4033")
darkDust = c("#1b0000")
darkForest = c("#06402B")

data <- read_csv("processedData/processedDirectorData.csv")
```

###### 

###### 

## Re-formatting Data

Just going to re-format the data into more intuitive names, because the way we collected data in Js is non-intuitive.

```{r}
selectedData <- data %>% 
  select(
    -list,
    -accurate, 
    -matLabel,
    -colLabel, 
    -accurate, 
    -non_filler_count, 
    -phase,
    -hex,
    -matCode
         ) %>% 
  
  mutate (redundant = recode(
      sufficient,
      color = "material",
      sound = "color"
      )) %>% 
  
  mutate (sufficient = recode(
    sufficient, 
    color = "color",
    sound = "material"
    )) %>% 
  
  mutate (displayType = recode(
    relevance,
    base = "baseline",
    relevant = "SLow_RHigh",
    irrelevant = "SHigh_RLow"
  )) %>% 
  mutate(
    displayType = if_else(trial == "filler", 
                          "baseline", 
                          displayType)
    ) %>%
  select(-relevance) %>%
  
  mutate (globalDiscriminability = recode(
    discriminability,
    high = "everythingHigh",
    low = "lowAttributePresent"
  )) %>%
  select(-discriminability) %>%
  
  mutate (overinformative = overdescription) %>% 
  select(-overdescription) 
```

```{r}
factoredData <- selectedData %>%
  mutate(
    overinformative = factor(overinformative, 
                             levels = c(TRUE, FALSE)), 
    displayType = factor(displayType, 
                       levels = c("SLow_RHigh", "SHigh_RLow", "baseline")),
    redundant = factor(redundant, 
                        levels = c("color", "material"))
    ) 
```

```{r}
factoredCriticalData <- factoredData %>% 
  filter(trial == "critical") 
```

## Null Model

*Set up null model with no predictors for model comparison*

```{r}
nullModel <- brm(
  formula = overinformative ~ 0 + Intercept + (1 | participant),
  data = factoredCriticalData,
  family = bernoulli(link = "logit"),
  warmup = 5000,   
  iter = 10000,     
  chains = 4,     
  cores=4,  
  prior = prior("normal(0, 2)", class = "b"),
  seed = 3000, 
  save_pars = save_pars(all = TRUE)
  )
```

###### 

###### 

```{r}
summary(nullModel)
```

## Display Type and Attribute-Redundant Predictors, with Interaction

### Visualisation

```{r}
labels <- c(
  "color" = "Colour Redundant",
  "material" = "Material Redundant"
)
  
selectedData %>% 
  filter(trial == "critical") %>% 
  mutate(od = ifelse(overinformative, 1, 0)) %>% 
  ggplot(
    aes(x = displayType, 
        y = od, 
        fill = displayType,
        color = displayType)
    ) + 
  facet_wrap(~redundant,
             labeller = as_labeller(labels)
               ) +
  stat_summary(fun.y = mean, geom="bar", 
               alpha = 0.9, show.legend=FALSE, 
               color = NA) + 
  geom_jitter(
      stat = "summary", aes(group = participant),
      fun = mean, alpha = 0.15, size = 2.8, 
      show.legend = FALSE, width = 0.1
      ) +
  stat_summary(fun.data = mean_cl_boot, 
               geom="linerange", size = 1, 
               color = "grey25") +
  coord_cartesian(ylim = c(0, 1)) +
  ylab("Proportion Overinformative") + 
  xlab("Display Type") +
  theme(
    text = element_text(size = 18, 
                        family = "serif",
                        color = "black"), 
    panel.border = element_rect(color = "gray80", 
                                fill = NA, 
                                size = 0.7),
    panel.grid = element_blank()
    ) +
  scale_x_discrete(
    limits = c("baseline", "SHigh_RLow", "SLow_RHigh"), 
    labels = c("Baseline", "S-High\nR-Low","S-Low\nR-High")
    ) + 
   scale_fill_manual(values = c("#e0cda9", "#9E8472", "#a0ae9b" )) +
  scale_color_manual(values = c("#5C4033", "#1b0000", "#06402B" )) 
```

###### 

### Run Model

```{r}
displayAttributeInteraction <- brm(
 formula = overinformative ~ 0 + Intercept + 
  redundant*displayType + (1 | participant),
  data = factoredCriticalData, 
  family = bernoulli(link = "logit"), 
  warmup = 5000,  
  iter = 10000,      
  chains = 4,       
  cores=4,   
  prior = prior("normal(0,2)", class = "b"),
  sample_prior = 'yes', 
  seed = 3000,       
  save_pars = save_pars(all = TRUE)
) 
```

###### 

### Model Summary

```{r}
stanplot(displayAttributeInteraction, 
         type = "trace")
stanplot(displayAttributeInteraction, 
               type = "acf_bar")
summary(displayAttributeInteraction)
```

###### 

###### 

## Display Type and Attribute-Redundant Predictors, no Interaction

### Run Model

```{r}
displayAttribute <- brm(
 formula = overinformative ~ 0 + Intercept + 
  redundant + displayType + (1 | participant),
  data = factoredCriticalData, 
  family = bernoulli(link = "logit"), 
  warmup = 5000,  
  iter = 10000,      
  chains = 4,       
  cores=4,   
  prior = prior("normal(0,2)", class = "b"),
  sample_prior = 'yes', 
  seed = 3000,       
  save_pars = save_pars(all = TRUE)
) 
```

###### 

### Model Summary

```{r}
stanplot(displayAttribute, 
         type = "trace")
stanplot(displayAttribute, 
               type = "acf_bar")
summary(displayAttribute)
```

###### 

###### 

## Attribute Only

#### Visual

```{r}
selectedData %>% 
  filter(trial == "critical") %>% 
  mutate(od = ifelse(overinformative, 1, 0)) %>% 
  ggplot(
    aes(x = redundant, 
        y = od, 
        fill = redundant,
        color = redundant)
    ) + 
  stat_summary(fun.y = mean, geom="bar", 
               alpha = 0.9, show.legend=FALSE, 
               color = NA) + 
  geom_jitter(
      stat = "summary", aes(group = participant),
      fun = mean, alpha = 0.2, size = 2.8, 
      show.legend = FALSE, width = 0.1
      ) +
  stat_summary(fun.data = mean_cl_boot, 
               geom="linerange", size = 1, 
               color = "grey25") +
  coord_cartesian(ylim = c(0, 1)) +
  ylab("Proportion Overinformative") + 
  xlab("Attribute Redundant") +
  theme(
    text = element_text(size = 18, 
                        family = "serif",
                        color = "black"), 
    panel.border = element_rect(color = "gray80", 
                                fill = NA, 
                                size = 0.7),
    panel.grid = element_blank()
    ) +
  scale_x_discrete(
    limits = c("color", "material"), 
    labels = c("Color Redundant", "Material Redundant")
    ) + 
   scale_fill_manual(values = c("#e0cda9", "#a0ae9b")) +
  scale_color_manual(values = c("#5C4033", "#06402B"))
```

###### 

#### Model

```{r}
attributeOnly <- brm(
 formula = overinformative ~ 0 + Intercept + 
  redundant + (1 | participant),
  data = factoredCriticalData, 
  family = bernoulli(link = "logit"), 
  warmup = 5000,  
  iter = 10000,      
  chains = 4,       
  cores=4,   
  prior = prior("normal(0,2)", class = "b"),
  sample_prior = 'yes', 
  seed = 3000,       
  save_pars = save_pars(all = TRUE)
) 
```

###### 

### Summary

```{r}
stanplot(attributeOnly, 
         type = "trace")
stanplot(attributeOnly, 
               type = "acf_bar")
summary(attributeOnly)
```

###### 

###### 

## Display Type Only

### Visual

```{r}
selectedData %>% 
  filter(trial == "critical") %>% 
  mutate(od = ifelse(overinformative, 1, 0)) %>% 
  ggplot(
    aes(x = displayType, 
        y = od, 
        fill = displayType,
        color = displayType)
    ) + 
  stat_summary(fun.y = mean, geom="bar", 
               alpha = 0.9, show.legend=FALSE, 
               color = NA) + 
  geom_jitter(
      stat = "summary", aes(group = participant),
      fun = mean, alpha = 0.2, size = 2.8, 
      show.legend = FALSE, width = 0.1
      ) +
  stat_summary(fun.data = mean_cl_boot, 
               geom="linerange", size = 1, 
               color = "grey25") +
  coord_cartesian(ylim = c(0, 1)) +
  ylab("Proportion Overinformative") + 
  xlab("Display Type") +
  theme(
    text = element_text(size = 18, 
                        family = "serif",
                        color = "black"), 
    panel.border = element_rect(color = "gray80", 
                                fill = NA, 
                                size = 0.7),
    panel.grid = element_blank()
    ) +
  scale_x_discrete(
    limits = c("baseline", "SHigh_RLow", "SLow_RHigh"), 
    labels = c("Baseline", "S-High\nR-Low", "S-Low\nR-High")
    ) + 
   scale_fill_manual(values = c("#e0cda9", "#a0ae9b", "#9E8472")) +
  scale_color_manual(values = c("#5C4033", "#06402B", "#1b0000")) 
```

###### 

### Model

```{r}
display <- brm(
 formula = overinformative ~ 0 + Intercept + 
  displayType + (1 | participant),
  data = factoredCriticalData, 
  family = bernoulli(link = "logit"), 
  warmup = 5000,  
  iter = 10000,      
  chains = 4,       
  cores=4,   
  prior = prior("normal(0,2)", class = "b"),
  sample_prior = 'yes', 
  seed = 3000,       
  save_pars = save_pars(all = TRUE)
) 
```

###### 

### Summary

```{r}
stanplot(display, 
         type = "trace")
stanplot(display, 
               type = "acf_bar")
summary(display)
```

###### 

###### 

## Global Discriminability Model

### Visual

```{r}
selectedData %>% 
  filter(trial == "critical") %>% 
  mutate(od = ifelse(overinformative, 1, 0)) %>% 
  ggplot(
    aes(x = globalDiscriminability, 
        y = od, 
        fill = globalDiscriminability,
        color = globalDiscriminability)
    ) + 
  stat_summary(fun.y = mean, geom="bar", 
               alpha = 0.9, show.legend=FALSE, 
               color = NA) + 
  geom_jitter(
      stat = "summary", aes(group = participant),
      fun = mean, alpha = 0.2, size = 2.8, 
      show.legend = FALSE, width = 0.1
      ) +
  stat_summary(fun.data = mean_cl_boot, 
               geom="linerange", size = 1, 
               color = "grey25") +
  coord_cartesian(ylim = c(0, 1)) +
  ylab("Proportion Overinformative") + 
  xlab("Global Discriminability") +
  theme(
    text = element_text(size = 18, 
                        family = "serif",
                        color = "black"), 
    panel.border = element_rect(color = "gray80", 
                                fill = NA, 
                                size = 0.7),
    panel.grid = element_blank()
    ) +
   scale_fill_manual(values = c("#e0cda9", "#a0ae9b")) +
  scale_color_manual(values = c("#5C4033", "#06402B"))
```

###### 

### Model

```{r}
globalDiscrimModel <- brm(
 formula = overinformative ~ 0 + Intercept + 
  globalDiscriminability + (1 | participant),
  data = factoredCriticalData, 
  family = bernoulli(link = "logit"), 
  warmup = 5000,  
  iter = 10000,      
  chains = 4,       
  cores=4,   
  prior = prior("normal(0,2)", class = "b"),
  sample_prior = 'yes', 
  seed = 3000,       
  save_pars = save_pars(all = TRUE)
) 
```

###### 

### Summary

```{r}
stanplot(globalDiscrimModel, 
         type = "trace")
stanplot(globalDiscrimModel, 
               type = "acf_bar")
summary(globalDiscrimModel)
```

###### 

###### 

## Scene Variation Model

### Visual

Filler vs. Critical models — whether scene variation increases overinformativeness. Low scene-variation is fillers, whereas high scene variation is criticals. Of course, we have to filter in ONLY high-discriminability trials to avoid confounding. This also means that this model can't be compared to the other models.

```{r}
selectedData %>% 
  mutate(od = ifelse(overinformative, 1, 0)) %>% 
  filter(globalDiscriminability == "everythingHigh") %>% 
  ggplot(
    aes(x = trial, 
        y = od, 
        fill = trial,
        color = trial)
    ) + 
  stat_summary(fun.y = mean, geom="bar", 
               alpha = 0.9, show.legend=FALSE, 
               color = NA) + 
  geom_jitter(
      stat = "summary", aes(group = participant),
      fun = mean, alpha = 0.15, size = 2.8, 
      show.legend = FALSE, width = 0.1
      ) +
  stat_summary(fun.data = mean_cl_boot, 
               geom="linerange", size = 1, 
               color = "grey25") +
  coord_cartesian(ylim = c(0, 1)) +
  ylab("Proportion Overinformative") + 
  xlab("Scene Variation") +
  theme(
    text = element_text(size = 18, 
                        family = "serif",
                        color = "black"), 
    panel.border = element_rect(color = "gray80", 
                                fill = NA, 
                                size = 0.7),
    panel.grid = element_blank()
    ) +
  scale_x_discrete(
    limits = c("critical", "filler"), 
    labels = c("High", "Low")
    ) +
  scale_fill_manual(values = c("#b37a5f", "#d4b88a")) + 
  scale_color_manual(values = c("#553929", "#68553d"))
```

Note that critical = high scene variation, filler = low scene variation.

###### 

### Model

```{r}
sceneVariationData <- selectedData %>%
  filter(globalDiscriminability == "everythingHigh")
```

```{r}
sceneVariation <- brm(
 formula = overinformative ~ 0 + Intercept + 
  trial + (1 | participant),
  data = selectedData, 
  family = bernoulli(link = "logit"), 
  warmup = 5000,  
  iter = 10000,      
  chains = 4,       
  cores=4,   
  prior = prior("normal(0,2)", class = "b"),
  sample_prior = 'yes', 
  seed = 3000,       
  save_pars = save_pars(all = TRUE)
) 
```

###### 

### Summary

```{r}
stanplot(sceneVariation, 
         type = "trace")
stanplot(sceneVariation, 
               type = "acf_bar")
summary(sceneVariation)
```

###### 

###### 

## Model Comparison: Evidence over the null hypothesis

```{r}
bayes_factor(displayAttributeInteraction, nullModel, log = TRUE)
bayes_factor(displayAttribute, nullModel, log = TRUE)
bayes_factor(attributeOnly, nullModel, log = TRUE)
bayes_factor(display, nullModel, log = TRUE)
bayes_factor(globalDiscrimModel, nullModel, log = TRUE)
```